
---
title: "Predicting Exercise Manner with Accelerometer Data"
author: "Wulan Margaretta"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal of this project is to predict the manner in which participants performed barbell lifts using data collected from accelerometers. The response variable, `classe`, indicates the exercise manner. Our task is to build a machine learning model to predict `classe` using the other variables, and to estimate the out-of-sample error.

## Data Preprocessing

```{r load-libraries}
library(caret)
library(randomForest)
library(dplyr)
```

```{r load-data}
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(train_url, na.strings = c("NA", "", "#DIV/0!"))
testing <- read.csv(test_url, na.strings = c("NA", "", "#DIV/0!"))

training <- training[, colSums(is.na(training)) == 0]
training <- training[, -c(1:7)]
testing <- testing[, colnames(testing) %in% colnames(training)]

testing <- testing[, names(training)[-length(training)]]
```

## Model Training

```{r model-training}
set.seed(1234)
inTrain <- createDataPartition(training$classe, p = 0.7, list = FALSE)
trainSet <- training[inTrain, ]
testSet <- training[-inTrain, ]

control <- trainControl(method = "cv", number = 5)
rfModel <- train(classe ~ ., data = trainSet, method = "rf", trControl = control)
```

## Model Evaluation

```{r model-evaluation}
rfPred <- predict(rfModel, testSet)
confMat <- confusionMatrix(rfPred, testSet$classe)
confMat
```

## Expected Out-of-Sample Error

```{r out-of-sample-error}
1 - confMat$overall['Accuracy']
```

## Final Predictions

```{r final-prediction}
finalPredictions <- predict(rfModel, testing)
finalPredictions
```

```{r write-files, eval=FALSE}
pml_write_files <- function(x){
  for(i in 1:length(x)){
    filename <- paste0("problem_id_", i, ".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}
pml_write_files(finalPredictions)
```

## Conclusion

We used a Random Forest model trained on accelerometer data to accurately predict exercise performance. Using cross-validation, we estimated a low out-of-sample error, indicating strong predictive performance.

## References

- Velloso, E. et al. (2013). *Qualitative Activity Recognition of Weight Lifting Exercises*. http://groupware.les.inf.puc-rio.br/har
